Dark Control: The Default Mode Network as a Reinforcement Learning Agent

### 2018-08-12 paper/paper.tex
gros commit

\section{Towards a formal account of default mode function: higher-order control of the organism}
%
In the absence of external stimulation, the human brain is not at rest.
At the turn to the 21st century,
brain-imaging may have been the first technique
to allow for the discovery of a unique brain network that would subserve
baseline mental activities
\citep{raichle2001pnas, randy2008, bzdok2015resting}.
The ``default mode network'' (DMN) continues
to metabolize large quantities of
oxygen and glucose energy to maintain
neuronal computation
during free-ranging thought
\citep{kenet2003spontaneously, fiser2004small}.
The baseline energy demand is only weakly modulated
at the onset of defined psychological tasks \citep{raichle_baseline}.
At its opposite, during sleep, the
decoupling of brain structures discarded the idea of the DMN being
only a passive network resonance and rather supported an
important role in sustaining conscious awareness \citep{horovitz2009decoupling}.

This \textit{dark matter of brain physiology}
\citep{raichle2006brain} begs the question
of the biological purpose underlying neural activity in the DMN,
which however still remains elusive at the electrophysiological level
\citep{de2010temporal,brookes2011investigating,baker2014fast}.
% the link between the fMRI BOLD signal and population-level neural activity is still unclear. If those frequency-specific electrophysiological correlations are proposed as complementary to those observed with BOLD \citep{hipp2015bold}, their role in DMN function remains elusive \citep{maldjian2014graph}.
What has early been described as the ``stream of consciousness''
in psychology \citep{james1890principles}
found a potential neurobiological manifestation
in the DMN
\citep{shul1997, raichle2001pnas}.
%
Axonal tracing injection in such parts of the association cortex in monkeys
were shown to resemble connectivity links between nodes of the human
DMN (see here for details on anatomical connections: \citep{randy2008}).
Additionally, myelination patterns of axon connections were found to finish
particularly late in these cortical areas \citep{flechsig1920}, often believed to suggest
sophistication of subserved neural processes
\citep{Sowell2003, Yakovlev1967}.
%
We propose that this set of some of the most advanced regions in
the association cortex \citep{mesulam1998sensation, margulies2016situating}
are responsible for higher-order control of human behavior.
Our functional account follows the notion of
``a hierarchy of brain systems with
the DMN at the top and the salience and dorsal attention systems
at intermediate levels, above thalamic and unimodal sensory
cortex'' \citep{carhart2010default}.

% \subsection{Towards a formal account of default mode function: higher-order control of the organism}
The network nodes that compose the human DMN
are hubs of high baseline neural activity,
which typically decreases when engaged in controlled psychological experiments
\citep{raichle_baseline}.
The standard mode of
neural information maintenance and manipulation has
been argued to mediate evolutionarily conserved functions
\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
Today, many psychologists and neuroscientists believe that
the DMN implements some form of
probabilistic estimation of past, hypothetical, and
future events
\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
randy2008, spreng2009common}.
This brain network
might have emerged to continuously predict the environment using
mental imagery as an evolutionary advantage \citep{suddendorf2007evolution}.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode regions \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Gaining insight into DMN function is
particularly challenging because
this brain network appears to
simultaneously modulate perception-action cycles in the present
and to support mental travel
across time, space, and content domains
\citep{boyer2008evolutionary}.

We aim at proposing an alternative to reasoning about the DMN
based on longstanding cognitive theory.
The present work adopts the control-theoretical perspective of
a human \textit{agent} faced with the choice of the next actions
and guided by outcomes to optimize behavioral performance. These outcomes
can be really experienced, hypothetically imagined, or simply expected in the future.
Formally, we propose reinforcement learning as a particularly attractive framework
for describing, containing, and quantifying the unknown function underlying DMN activity.
An intelligent agent improves the interaction with the environment
by continuously updating its computation of value estimates and action predispositions
through integration of feedback outcomes.
That is, ``[agents], with their actions, modify the environment and in doing so
partially determine their next stimuli, in particular stimuli that are necessary
for triggering the next action''
\citep{pezzulo2011grounding}.
Agents with other behavioral policies therefore sample different
distributions of action-perception trajectories \citep{ghavamzadeh2015bayesian}.
Henceforth,
\textit{control} refers to the influence that an agent exerts when interacting
with the environment to reach preferred states.

@@

Such a RL perspective on DMN function can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term gratification.
We argue that DMN implication in many
of the most advanced human capacities
can be recast as prediction error minimization
informed by internally generated probabilistic simulations
- ``covert forms of action and perception'' \citep{pezzulo2011grounding} -,
allowing maximization of action outcomes across different time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Monte Carlo sampling.
Even necessarily imperfect memory recall,
random day-time mind-wandering, and
seemingly arbitrary dreams during sleep
may provide randomly sampled blocks of pseudo-experience  that are
instrumental to iteratively
optimize the behavioral agenda of the organism.

@@

Please appreciate the importance of differentiating which levels of observation are at play in the present account. A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent. Henceforth, we will use ``inference`` when referring to aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
%
It is moreover important to note that
our account does not claim that neural activity in the DMN in particular
or the brain in general are identical with reinforcement learning algorithms.
Rather, we advocate feedback-based learning strategies as an attractive alternative perspective
to describe, quantify, and interpret research findings on the DMN.

@@

  \includegraphics[width=.75\linewidth]{fig_dmn.png}
  \caption{\textbf{Predictive structural association between reward system and
      DMN nodes.} Reward tasks \citep{doherty2015structure} and neural processing in the DMN \citep{randy2008}, often considered task-negative, have been studied so far in largely separate niches of the neuroscience literature. A currently underappreciated link is however suggested here based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$ [standard deviation across cross-validation folds]) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from (z-scored) volume in the DMN regions. These out-of-sample generalizations reflect the expected performance in yet-to-be observed individuals obtained from linear support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure \citep{hastie}.
    Consistent for the left and right reward system, NAc volume in a given subject is
    positively coupled with the vmPFC and HC. The congruence of our structural association results for both NAc targets speaks to the robustness of our pattern-prediction findings. The opposite relation of the left and right TPJ to the NAc appears to reflect a repeatedly recognized hemispheric asymmetry with respect to functional implications \citep{seghier2013angular}, impairments in neurological patients \citep{corbetta2000}, different types of connectivity \citep{uddin_tpj, caspers2011} as well as micro- and macroanatomy \citep{caspers2006, caspers2008}. The colors are indicative of the (\textcolor{red}{red} = positive,
    \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
  The code for reproduction and visualization: \url{www.github.com/banilo/darkcontrol_2018}.}
  \label{fig:dmn}
% \end{wrapfigure}
\end{figure}

@@

\subsection{The right and left TPJ: prediction error signaling and world semantics}
The DMN emerges with its midline structures early in human development
\citep{doria2010}, while
the right and left TPJs may become fully functionally integrated into this major brain
network only after birth.
The TPJs are known to exhibit hemispheric differences
based on microanatomical properties and cortical gyrification patterns
\citep{seghier2013angular}.
In general, neuroscientific investigations on hemispheric functional specialization
have highlighted the right cerebral hemisphere as more dominant for attentional functions
and the left side more for semantic functions
\citep{seghier2013angular, bzdok2013tpj, bzdok2016left,
stephan2007mechanisms}.

The TPJ in the right hemisphere (RTPJ) denotes a broad functional zone with
varying anatomical nomenclature \citep{marstpj2011, seghier2010, seghier2013angular}
that has been shown to be closely related to
multi-sensory prediction and prediction error signaling
\citep{downar2000multimodal, vetter2011, shulman2010, shulman2007}.
This DMN region is probably central for
action initiation during goal-directed psychological tasks and for
sensorimotor behavior by integrating multi-sensory attention
\citep{corbetta2002control}.
Its involvement was repeatedly reported in
multi-step action execution \citep{hartmann2005takes},
visuo-proprioceptive conflict \citep{Balslev2005},
spatial reorienting \citep{corbetta2000},
and
detection of environmental changes across
visual, auditory, or tactile stimulation
\citep{downar2000multimodal}.
Direct electrical stimulation of the human
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness \citep{blanke2002neuropsychology}.
%
It was argued that the RTPJ encodes actions and predicted outcomes,
without necessarily relating these neural processes to value estimation
\citep{liljeholm2013neural, hamilton2008action,
jakobs2009effects}.
More specifically, neural activity in the RTPJ has been proposed to reflect
stimulus-driven attentional reallocation to
self-relevant and unexpected sources of information
as a ‘‘circuit breaker’’ that recalibrates functional control of brain networks
\citep{bzdok2013tpj, corbettashul2008}.
% (Bzdok et al., 2013; Corbetta et al., 2008).
% Indeed, neurological patients with RTPJ damage have particular difficulties
% with multi-step actions \citep{hartmann2005takes}.
In the face of large discrepancies between actual and previously predicted
environmental events, the RTPJ acts as a potential switch between
externally-oriented mind sets focussed on the
sensory environment and internally-oriented mind sets focussed
on mental scene construction.
For instance, temporally induced RTPJ damage in humans diminished the
impact of predicted intentions of other individuals
\citep{young2010disruption},
a capacity believed to be enabled by the DMN.
Viewed from a RL account,
the RTPJ might reflect an important relay that shifts away
from the ‘‘internally directed’’ baseline processes
to, instead, deal with unexpected environmental stimuli and events.
% (TODO: slow endogeneous changes in state transitions by the policy and value matrices)

@@

\section{Reinforcement learning control: a process model for DMN function}
We argue the outlined neurobiological properties
of the DMN regions
to be sufficient for implementing all components
of a full-fledged reinforcement learning (RL) system.
Recalling past experience, considering candidate actions,
random sampling of possible experiences, as well as
estimation of instantaneous and expected delayed reward outcomes
are key components of intelligent RL agents
that are plausible to functionally intersect in the DMN.

RL is an area of machine learning concerned with searching optimal behavioral strategies through interactions
with an \textit{environment} with the goal to maximize the \textit{cumulative reward} over time.
The optimal behavior typically
takes the future into account as certain rewards could be \textit{delayed}.
Through repeated action on and feedback from the environment,
the agent learns how to reach goals and continuously improve the collection of reward signals
in a trial-and-error fashion (Fig. \ref{fig:rl}).
At a given moment, each taken \textit{action} $a$ triggers a change
in the \textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
$r = r(s, a,s')$ obtained by the agent.
If the collected reward outcome yields a negative value it can be
more naturally interpreted as \textit{punishment}.
The environment be partly controlled by
the action of the agent and the reward can be thought
of as satisfaction \textemdash or aversion \textemdash that accompany the execution of
a particular action.
% Action may be delayed to achieve substantial increases in expected reward
% that is expected to grow with time.

The environment is generally taken as \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory input
\citep{starkweather2017dopamine}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that an abstract description of DMN activity based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.
In short, DMN implication in a diversity of
particularly sophisticated human behaviors
can be parsimoniously explained as instantiating probabilistic
simulations of experience coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time scales.
Such a purposeful optimization objective
may be subserved by a stochastic approximation
based on a brain implementation for Monte Carlo sampling of events and outcomes.

@@

\subsection{Markov decision processes}
% RL has had considerable success in modeling complex decision-making problems,
% We emphasize that the brain in general, and the DMN in particular,
% is a physical system governed by the laws of
% physics and can be formally described
% by Markov processes at a sufficiently coarse scale.
% It has indeed been previously proposed \citep{tegmark2016improved} that
% any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
% step is sufficiently short.
In artificial intelligence and machine learning, a popular computational model for
multi-step decision processes are
MDPs \citep{sutton1998reinforcement}.
% For the remainder of this paper, for the sake of simplicity
% we use the abbreviation ``MDP'' to refer to the special case
% of \textit{Partially Observable Markov Decision Processes}.
An MDP operationalizes a sequential decision process
in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state.
Instead, the agent tries to optimize a \textit{subjective} reward
signal (i.e., likely to be different for another agent in the same state
and possibly driven by neural processing in the vmPFC)
by maintaining probability distributions over actions according to their expected utility.
% This is performed alongside each transition in the state space of the environment
%  and maintains a probability distribution over the set of possible
% states.
% % , based on a set of observations and observation probabilities.
This is a minimal set of assumptions that can be made about an environment faced by an agent engaged in interactive learning.

\begin{definition*}
Mathematically, an MDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p)$ where
\begin{itemize}
\item $\mathcal S$ is the set of states, such as $\mathcal S = \{\text{happy}, \text{sad}, \text{puzzeled}\}$.
\item $\mathcal A$ is the set of actions, such as $\mathcal A = \{\text{read}, \text{run},
  \text{laugh}, \text{sympathize}, \text{empathize}\}.$
\item $r : \mathcal S \times \mathcal A \times \mathcal S \rightarrow \mathbb R$ is the \textit{reward function},
   so that $r(s, a, s')$ is the instant reward for taking action $a$ in state $s$ followed by a state-transition $s \rightarrow s'$.
% \item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
%   $\mu(s)$ is the probability that the environment starts off in state $s$.
\item $p : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto p(s'|s,a)$,
  the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present state and action taken.
\end{itemize}
\end{definition*}

% , defining $s(t)$, the state at time $t$, as the position in state space.
The process has $memory$ if the subsequent state depends not only on the current state but also on a number of past states. Rational probabilistic planning can thus be reformulated as a standard memoryless Markov process by simply expanding the definition of the state $s$ to include experience episodes of the past. This extension adds the capacity for memory to the model because the next state then depends not only on the current situation but also on previously experienced events, which is the motivation behind Partially Observable MDPs (POMDPs)
\citep{starkweather2017dopamine,o2006making}.
Nevertheless, this mathematical property of POMDPs mostly accounts for implicit $memory$. Since the current paper is concerned with plausibility at the behavioral and neurobiological level, we will address below how our account can accommodate the neurophysiological constraints of the DMN and the explicit $memory$ characteristics of human agents.

\paragraph*{Why Markov Decision Processes?} One may wonder whether MDP models
are applicable to something as complex as human behavior. This class of
reinforcement-learning models has had numerous successes in diverse domains. For instance, financial trading is largely a manifestation of strategic decision-making of interacting human agents. According to how the market responds, the agent incurs gain or loss as environmental feedback of the executed financial actions. Recent research on automatizing market exchanges by algorithmic trading has effectively deployed MDPs as a framework for modeling these elaborate behavioral dynamics  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010, abergel2017}.
%
MDPs have also been effective as a behavioral model in robotics
\citep{ng2004,abbeel2004} and in challenging multistep strategy games
~\citep{mnih2015,silver2016mastering,pritzel2017neural}. 
More recent work has
developped an MDP-based way of reasoning about future behavior
of other agents \citep{robinowitz2018}.
The idea is to use meta-learning (i.e., learning to learn) to build strong priors
about the behavior of a population of other agents.

>>>>

\section{Introduction}
%
In the absence of external stimulation, the human brain is not at rest.
At the turn to the 21st century,
brain-imaging may have been the first technique
to allow for the discovery of a unique brain network that would subserve
baseline mental activities
\citep{raichle2001pnas, randy2008, bzdok2015resting}.
The ``default mode network'' (DMN) continues
to metabolize large quantities of
oxygen and glucose energy to maintain
neuronal computation
during free-ranging thought
\citep{kenet2003spontaneously, fiser2004small}.
The baseline energy demand is only weakly modulated
at the onset of defined psychological tasks \citep{raichle_baseline}.
At its opposite, during sleep, the
decoupling of brain structures discarded the idea of the DMN being
only a passive network resonance and rather supported an
important role in sustaining conscious awareness \citep{horovitz2009decoupling}.

This \textit{dark matter of brain physiology}
\citep{raichle2006brain} begs the question
of the biological purpose underlying neural activity in the DMN,
which however still remains elusive at the electrophysiological level
\citep{de2010temporal,brookes2011investigating,baker2014fast}.
% the link between the fMRI BOLD signal and population-level neural activity is still unclear. If those frequency-specific electrophysiological correlations are proposed as complementary to those observed with BOLD \citep{hipp2015bold}, their role in DMN function remains elusive \citep{maldjian2014graph}.
What has early been described as the ``stream of consciousness''
in psychology \citep{james1890principles}
found a potential neurobiological manifestation
in the DMN
\citep{shul1997, raichle2001pnas}.
%
Axonal tracing injection in such parts of the association cortex in monkeys
were shown to resemble connectivity links between nodes of the human
DMN (see here for details on anatomical connections: \citep{randy2008}).
Additionally, myelination patterns of axon connections were found to complete
particularly late in these cortical areas \citep{flechsig1920}, often believed to reflect
sophistication of subserved neural processes
\citep{Sowell2003, Yakovlev1967}.
%
We propose that this set of some of the most advanced regions in
the association cortex \citep{mesulam1998sensation, margulies2016situating}
are responsible for higher-order control of human behavior.
Our perspective therefore follows the notion of
``a hierarchy of brain systems with
the DMN at the top and the salience and dorsal attention systems
at intermediate levels, above thalamic and unimodal sensory
cortex'' \citep{carhart2010default}.

\section{Towards a formal account of default mode function: higher-order control of the organism}
The network nodes that compose the human DMN
are hubs of high baseline neural activity,
which typically decreases when engaged in well-defined psychological experiments
\citep{raichle_baseline}.
The standard mode of
neural information maintenance and manipulation has
been argued to mediate evolutionarily conserved functions
\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
Today, many psychologists and neuroscientists believe that
the DMN implements some form of
probabilistic estimation of past, hypothetical, and
future events
\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
randy2008, spreng2009common}.
This brain network
might have emerged to continuously predict the environment using
mental imagery as an evolutionary advantage \citep{suddendorf2007evolution}.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode regions \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Gaining insight into DMN function is
particularly challenging because
this brain network appears to
simultaneously influence perception-action cycles in the present
and to support mental travel
across time, space, and content domains
\citep{boyer2008evolutionary}.

We aim at proposing an alternative to reasoning about the DMN
based on longstanding cognitive theory.
The present work adopts the control-theoretical perspective of
a human \textit{agent} faced with the choice of the next actions
guided by outcomes to optimize behavioral performance. These outcomes
can be really experienced, hypothetically imagined, or expected in the future.
Formally, we propose reinforcement learning as a particularly attractive framework
for describing, containing, and quantifying the unknown function underlying DMN activity.
An intelligent agent improves the interaction with the environment
by continuously updating its computation of value estimates and action predispositions
through integration of feedback outcomes.
That is, ``[agents], with their actions, modify the environment and in doing so
partially determine their next stimuli, in particular stimuli that are necessary
for triggering the next action''
\citep{pezzulo2011grounding}.
Agents with other behavioral policies therefore sample different
distributions of action-perception trajectories \citep{ghavamzadeh2015bayesian}.
Henceforth,
\textit{control} refers to the influence that an agent exerts by interacting
with the environment to reach preferred states.

@@

Such a RL perspective on DMN activity can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term gratification.
We argue that DMN implication in many
of the most advanced human capacities
can be recast as prediction error minimization
informed by internally generated probabilistic simulations
- ``covert forms of action and perception'' \citep{pezzulo2011grounding} -,
allowing maximization of action outcomes across different time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Monte Carlo sampling.
Even necessarily imperfect memory recall,
random day-time mind-wandering, and
seemingly arbitrary dreams during sleep
may provide randomly sampled blocks of pseudo-experience  that are
instrumental to iteratively
optimize the behavioral agenda of the organism.

@@

Please appreciate the importance of differentiating which levels of observation are at play in the present account. A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent. Henceforth, we will use ``inference`` when referring to aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the
cognitive behavior of the agent.
%
It is moreover important to note that
our account does not claim that neural activity in the DMN in particular
or the brain in general are identical with reinforcement learning algorithms.
Rather, we advocate feedback-based learning strategies as an attractive alternative perspective
to describe, quantify, and interpret research findings on the DMN.

@@

  \includegraphics[width=.75\linewidth]{fig_dmn.png}
  \caption{\textbf{Predictive structural association between reward system and
      DMN nodes.} Reward tasks \citep{doherty2015structure} and neural processing in the DMN \citep{randy2008}, often considered ``task-negative'', have been studied so far in largely separate niches of the neuroscience literature. A currently underappreciated link is however suggested here based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$ [standard deviation across cross-validation folds]) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from (z-scored) volume in the DMN regions. These out-of-sample generalizations reflect the expected performance in yet-to-be observed individuals obtained from linear support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure \citep{hastie}.
    Consistent for the left and right reward system, NAc volume in a given subject is
    positively coupled with the vmPFC and HC. The congruence of our structural association results for both NAc targets speaks to the robustness of our pattern-prediction findings. The opposite relation of the left and right TPJ to the NAc appears to reflect a repeatedly recognized hemispheric asymmetry with respect to functional implications \citep{seghier2013angular}, impairments in neurological patients \citep{corbetta2000}, different types of connectivity \citep{uddin_tpj, caspers2011} as well as micro- and macroanatomy \citep{caspers2006, caspers2008}. The colors are indicative of the (\textcolor{red}{red} = positive,
    \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
  The code for reproduction and visualization: \url{www.github.com/banilo/darkcontrol_2018}.}
  \label{fig:dmn}
% \end{wrapfigure}
\end{figure}

@@

\subsection{The right and left TPJ: prediction error signaling and world semantics}
The DMN emerges with its midline structures early in human development
\citep{doria2010}, while
the right and left TPJs may become fully functionally integrated into this macroscopical
network only after birth.
The TPJs are known to exhibit hemispheric differences
based on microanatomical properties and cortical gyrification patterns
\citep{seghier2013angular}.
In general, neuroscientific investigations on hemispheric functional specialization
have highlighted the right cerebral hemisphere as more dominant for attentional functions
and the left side more for semantic functions
\citep{seghier2013angular, bzdok2013tpj, bzdok2016left,
stephan2007mechanisms}.

The TPJ in the right hemisphere (RTPJ) denotes a broad functional zone with
varying anatomical nomenclature \citep{marstpj2011, seghier2010, seghier2013angular}
that has been shown to be closely related to
multi-sensory event representation and prediction error signaling
\citep{downar2000multimodal, vetter2011, shulman2010, shulman2007}.
This DMN region is probably central for
action initiation during goal-directed psychological tasks and for
sensorimotor behavior by integrating multi-sensory attention
\citep{corbetta2002control}.
Its involvement was repeatedly reported in monitoring
multi-step action execution \citep{hartmann2005takes},
visuo-proprioceptive conflict \citep{Balslev2005},
spatial re-orientation \citep{corbetta2000},
and
detection of environmental changes across
visual, auditory, or tactile stimulation
\citep{downar2000multimodal}.
Direct electrical stimulation of the human
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness \citep{blanke2002neuropsychology}.
%
It was argued that the RTPJ encodes actions and predicted outcomes,
without necessarily relating these neural processes to value estimation
\citep{liljeholm2013neural, hamilton2008action,
jakobs2009effects}.
More specifically, neural activity in the RTPJ has been proposed to reflect
stimulus-driven attentional reallocation to
self-relevant and unexpected sources of information
as a ‘‘circuit breaker’’ that recalibrates functional control of brain networks
\citep{bzdok2013tpj, corbettashul2008}.
% (Bzdok et al., 2013; Corbetta et al., 2008).
% Indeed, neurological patients with RTPJ damage have particular difficulties
% with multi-step actions \citep{hartmann2005takes}.
In the face of large discrepancies between actual and previously predicted
environmental events, the RTPJ acts as a potential switch between
externally-oriented mind sets focussed on the
sensory environment and internally-oriented mind sets focussed
on mental scene construction.
For instance, temporally induced RTPJ damage in humans diminished the
impact of predicted intentions of other individuals
\citep{young2010disruption},
a capacity believed to be enabled by the DMN.
Viewed from a RL perspective,
the RTPJ might reflect an important relay that shifts away
from the ‘‘internally directed’’ baseline processes
to, instead, deal with unexpected environmental cues and events.
% (TODO: slow endogeneous changes in state transitions by the policy and value matrices)

@@

\section{Reinforcement learning control: a process model for DMN function}
We argue the outlined neurobiological properties
of the DMN regions
to be sufficient for implementing all components
of a full-fledged reinforcement learning (RL) system.
Recalling past experience, considering candidate actions,
random sampling of possible experiences, as well as
estimation of instantaneous and delayed reward outcomes
are key components of intelligent RL agents
that are plausible to functionally intersect in the DMN.

RL is an area of machine learning concerned with searching optimal behavioral strategies through interactions
with an \textit{environment} with the goal to maximize the \textit{cumulative reward} over time
\citep{sutton1998reinforcement}.
Optimal behavior typically
takes the future into account as certain rewards could be \textit{delayed}.
Through repeated action on and feedback from the environment,
the agent learns how to reach goals and continuously improve the collection of reward signals
in a trial-and-error fashion (Fig. \ref{fig:rl}).
At a given moment, each taken \textit{action} $a$ triggers a change
in the \textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
$r = r(s, a,s')$ obtained by the agent.
If the collected reward outcome yields a negative value it can be
more naturally interpreted as \textit{punishment}.
The environment can be partly controlled by
the action of the agent and the reward can be thought
of as satisfaction \textemdash or aversion \textemdash that accompany the execution of
a particular action.
% Action may be delayed to achieve substantial increases in expected reward
% that is expected to grow with time.

The environment is assumed to be \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory perception
\citep{starkweather2017dopamine}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that an abstract description of DMN activity based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.
In short, DMN implication in a diversity of
particularly sophisticated human behaviors
can be parsimoniously explained as instantiating probabilistic
simulations of experience coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time scales.
Such a purposeful optimization objective
may be subserved by a stochastic approximation
based on a brain implementation for Monte Carlo sampling of events and outcomes.

@@

\subsection{Markov Decision Processes}
% RL has had considerable success in modeling complex decision-making problems,
% We emphasize that the brain in general, and the DMN in particular,
% is a physical system governed by the laws of
% physics and can be formally described
% by Markov processes at a sufficiently coarse scale.
% It has indeed been previously proposed \citep{tegmark2016improved} that
% any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
% step is sufficiently short.
In artificial intelligence and machine learning, a popular computational model for
multi-step decision processes are
MDPs \citep{sutton1998reinforcement}.
% For the remainder of this paper, for the sake of simplicity
% we use the abbreviation ``MDP'' to refer to the special case
% of \textit{Partially Observable Markov Decision Processes}.
An MDP operationalizes a sequential decision process
in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state.
Instead, the agent tries to optimize a \textit{subjective} reward
signal (i.e., likely to be different for another agent in the same state
and possibly driven by neural processing in the vmPFC)
by maintaining probability distributions over actions
(possibly represented in the dmPFC)
according to their expected utility.
% This is performed alongside each transition in the state space of the environment
%  and maintains a probability distribution over the set of possible
% states.
% % , based on a set of observations and observation probabilities.
This is a minimal set of assumptions that can be made about an environment faced by an agent engaged in interactive learning.

\begin{definition*}
Mathematically, an MDP involves a quadruple $(\mathcal S, \mathcal A, r, p)$ where
\begin{itemize}
\item $\mathcal S$ is the set of states, such as $\mathcal S = \{\text{happy}, \text{sad}, \text{puzzeled}\}$.
\item $\mathcal A$ is the set of actions, such as $\mathcal A = \{\text{read}, \text{run},
  \text{laugh}, \text{sympathize}, \text{empathize}\}.$
\item $r : \mathcal S \times \mathcal A \times \mathcal S \rightarrow \mathbb R$ is the \textit{reward function},
   so that $r(s, a, s')$ is the instant reward for taking action $a$ in state $s$ followed by a state-transition $s \rightarrow s'$.
% \item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
%   $\mu(s)$ is the probability that the environment starts off in state $s$.
\item $p : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto p(s'|s,a)$,
  the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present state and action taken.
\end{itemize}
\end{definition*}

% , defining $s(t)$, the state at time $t$, as the position in state space.
The process has $memory$ if the subsequent state depends not only on the current state but also on a number of past states. Rational probabilistic planning can thus be reformulated as a standard memoryless Markov process by simply expanding the definition of the state $s$ to include experience episodes of the past. This extension adds the capacity for memory to the model because the next state then depends not only on the current situation but also on previously experienced events, which is the motivation behind Partially Observable MDPs (POMDPs)
\citep{starkweather2017dopamine,o2006making}.
Nevertheless, this mathematical property of POMDPs mostly accounts for implicit memory. Since the current paper is concerned with plausibility at the behavioral and neurobiological level, we will address below how our account can accommodate the neurophysiological constraints of the DMN and the explicit memory characteristics of human agents.

\paragraph*{Why Markov Decision Processes?} One may wonder whether MDP models
are applicable to something as complex as human behavior. This class of
reinforcement-learning models has had numerous successes in diverse applied domains. For instance, MDPs have been successfully used in financial trading is largely a manifestation of strategic decision-making of interacting human agents. According to how the market responds, the agent incurs gain or loss as environmental feedback of the executed financial actions. Recent research on automatizing market exchanges by algorithmic trading has effectively deployed MDPs as a framework for modeling these elaborate behavioral dynamics  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010, abergel2017}.
%
MDPs have also been effective as a behavioral model in robotics
\citep{ng2004,abbeel2004} and in challenging multistep strategy games
~\citep{mnih2015,silver2016mastering,pritzel2017neural}. 
More recent work has
developped an MDP-related way of reasoning about future behavior
of other agents \citep{robinowitz2018}.
The idea is to use meta-learning (i.e., learning to learn) to build strong priors
about the behavior of a population of other agents.
